---
title: "Covid91 vaccine study FinalFall2022"
author: "Kathleen Castro"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalData <-read.csv(file="finalFall2022.csv",header=TRUE)
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
M = filter(FinalData, sex =="M")
F = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

#Graphical Descriptive Results
```{r}
dd2 <- M %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The graph represented above shows the population of Males in contact with Covid 19 through treatments on either placebo or vaccine. The graph represented above shows more than 60% of men who took placebo got Covid-19 while roughly 30% of men who took the vaccine got covid-19. This difference we see shows how the vaccine has become effective against the disease and gives a positive effect towards men.
The data presented above gives two graphs, one to represent the total count of infected and the other representing the percentage. The percentage looks to be more effected as it gives a more clear result while the total count gives a small data graph making it barely visable to know the difference.

#Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=M)
rowPerc(table1)
colPerc(table1)
```
  
The data represented above gives an overview on the specific percentages of men that got Covid-19 after taking either placebo or the vaccine. Roughly 31.56% of men got covid after taking the vaccine leaving roughly 68.44% of men who had taken placebo get covid. The vaccine looks like it is very effective and works very well for men.

#Inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
Mfish <- fisher.test(table1)
Mfish
```
  
The fisher test is an important test that provided us with the necessary p-value. The p-value recorded is 3e-07. The number turns out to be a very small amount which means that these results are more unlikely to occur by chance. The odds ratio is also given and is recorded to be 0.21.


```{r}
effect <- Mfish$p.value/(Mfish$p.value +1)
effectivenessM = Mfish$estimate/(Mfish$estimate +1)*100
print(effectivenessM)
```

```{r}
low=(Mfish$conf.int[1]/(Mfish$conf.int[1] +1))*100
hi=(Mfish$conf.int[2]/(Mfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The effectiveness is now used for the covid-19 vaccine within Males. If the result comes out to be above 50%, that means the vaccine becomes effective. Looking at our result, we can conclude that the vaccine is effective for men when trying not to contract covid-19. The odds ratio can also show how effective the vaccine is. Our collected odds ratio is 68.53436 which shows promising results as it reaches to a higher number.


## Females

#Barplot
```{r}
dd2 <- F %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The graph's represented above now show the comparison between the two drug types and getting covid-19 afterwords, but with women instead of men. Looking at the data given above, we can see that more women who have taken the vaccine end up with covid-19. It ranges to approximately 50%. Now taking a look at women who have taken placebo, we can see that there are fewer women who have gotten covid, ranging to approximately 42%. After reviewing this data set, we can conclude that women are more at risk of getting covid after recieving the vaccine compared to women taking placebo.


#Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=F)
rowPerc(table1)
colPerc(table1)
```
  
The numerical results shows specific percentages for females that received covid-19 after taking either drug. The females who had taken the vaccine and obtained covid was 57.18%. This shows how ineffective the vaccine is for women. The females who had taken Placebo and obtained covid-19 was 42.82%. This data shows how women are still at a high risk of getting covid-19 when taking this drug; however chances of getting covid are higher when getting the vaccine.


#Inferential
```{r}
chisq.test(table1)
chisqtestGC(table1)
Ffish <- fisher.test(table1)
Ffish
```
  
The fishers test shown above provided the required p-value which is 3e-07. This number came out to be a very small amount which proves that these results are unlikely to occur by chance. We can also see that the null hypothesis was rejected in favor of the alternative. Our recorded odds ratio is at 0.21.


```{r}
effect <- Ffish$p.value/(Ffish$p.value +1)
effectivenessF = Ffish$estimate/(Ffish$estimate +1)*100
print(effectivenessF)
```

```{r}
low=(Ffish$conf.int[1]/(Ffish$conf.int[1] +1))*100
hi=(Ffish$conf.int[2]/(Ffish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

  
If the effectiveness is above 50%, than the vaccine works, however for our data collected for the women, we can see that our effectiveness data is below 50%. This proves that the vaccine was ineffective.

## LGBTQ
###Barplot
```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The new set of data provided above now shows which drug taken will give the specific group, in this case the LGTBQ community, covid-19. The data shows that the gay's who had taken placebo have a lower chance of getting covid-19 afterwords. They range around 15%. The gay's who had taken the vaccine are at a higher amount, about 80%, of getting covid afterwords. This data shows that the vaccine isn't effective for members part of the LGBTQ community as more people receive covid when taking the vaccine


###Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```
  
Now reviewing the numerical results provided data, we can see the specific percentages for each drug. The amount of gay's who had taken the vaccine and obtained covid-19 after is at 86.67%. The amount of gay's who had taken placebo and obtained covid-19 after is at 13.33%. We can conclude that the drug placebo is more effective against Covid for the LGBTQ community compared to the vaccine which shows no effect. The high value proves that the vaccine doesn't help at all.

###Interential
```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```
  
The concluded p-value is 1.244e-08. This number comes out to a very small value showing that the results concluded are highly unlikely to occur by chance. The Fisher's Test is used to express the odds for Gay individuals which is 0.1617066.

```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgays = gaysfish$estimate/(gaysfish$estimate +1)*100
print(effectivenessgays)
```
```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))*100
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`
  
As stated before, if the effectiveness is above 50%, the vaccine is in effect; however our data resulted with an amount far below 50% showing that the vaccine does little to no effect for the LGBTQ community. This data proves that the vaccine won't work against covid for this group and is advise to avoid taking the vaccine. The odds ratio came to a very low number, 17, also showing the negative attributes of the vaccine.


## Druggies
###Barplot
```{r}
dd2 <-  Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The graph represented above gives data on the population of people who answered a poll on using drugs and taking the two discussed drugs. People who uses drugs that had taken placebo have a lower chance of receiving covid-19 ranging to around 18%. Drug users who have taken the vaccine are at a higher risk of getting covid-19 ranging over 80%. This data shown here proves that the drug users are not getting any positive effects after taking the vaccine compared to drug users who are taking placebo.


###Numerical Results
```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```
  
Taking a look at the numerical results, we can see that the data providing the specific percentages. Drug users are at an 83% chance of getting covid after taking the vaccine. Drug users who have taken placebo and had gotten covid are at a 17%. It appears the vaccine is ineffective to people in use of other drugs proving it to become useless for this specific population.

###Interential
```{r}
chisq.test(table1)
chisqtestGC(table1)
drugfish <- fisher.test(table1)
drugfish
```
  
The expressed p-value is at 3e-07. The number is a very small value proving that the results are unlikely to happen by chance. The Fisher test was able to reveal the odds ratio which is 0.21.


```{r}
effect <- drugfish$p.value/(drugfish$p.value +1)
effectivenessdrug = drugfish$estimate/(drugfish$estimate +1)*100
print(effectivenessdrug)
```

```{r}
low=(drugfish$conf.int[1]/(drugfish$conf.int[1] +1))*100
hi=(drugfish$conf.int[2]/(drugfish$conf.int[2] +1))*100
options(digits=2)
```

## Effectiveness Confidence Interval `r low`  to `r hi`

The results provided by the effectiveness came out to a very low number, below the 50% mark. We can With this data we can see how ineffective the vaccine is for drug users. The odds ratio also came out to be a low number, 17, proving how useless the vaccine has become for drug users.


# Overall Results and Conclusions
  
Looking over our data, we can conclude that the vaccine only works for one out of four of our population groups. It seems that men earn a more positive effect when taking the vaccine without having a high chance of getting covid-19 compared to women, LGBTQ, and drug users. Using the two graphs analysis, we can see the comparison of men using placebo and the vaccine when getting covid afterwords. The graphs for the other three groups show a complete opposite result where the vaccine give more negative results while the placebo's results look promising.
  After the graphs, we can see the numerical results which gives us a more accurate answer on the percentages. The concluded information for men gives the vaccine a more welcoming amount with excellent results where fewer men get covid after taking the vaccine. The other groups are different as they show the drug Placebo to be more promising than the vaccine.
    Now adding the fisher test, odds ratio, and the effectiveness analysis, we can collect more data on wither the vaccine is better than placebo or the other way around. We can see that for men, the vaccine shows great results but looking at the other groups, we can see that the results came out to be bad. With our collected data we can conclude that the covid-vaccine should not be used as it only positively effects men and does not help in other targeted groups. 


